Estimating Correspondences of Deformable Objects ``In-the-wild"

### 2016-01-27 rebuttal/egrebuttal.tex

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr_rebuttal}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

%%%%%%%%% PAPER ID  - PLEASE UPDATE
\def\cvprPaperID{0008} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
\setlength{\parindent}{0pt}

\usepackage{caption}

\begin{document}

%%%%%%%%% TITLE - PLEASE UPDATE
\title{Estimating Correspondences of Deformable Objects ``In-the-wild"}  % **** Enter the paper title here

\maketitle
\thispagestyle{empty}

%%%%%%%%% BODY TEXT - ENTER YOUR RESPONSE BELOW
We thank all Reviewers for their valuable feedback. Both Reviewers 24 and 25 (R24, R25) accept the paper (R24 strongly accepts it by acknowledging its potential impact). However, Reviewer 23 (R23) is reserved, but in the same time thinks it is ``interesting and quite novel''. It goes without saying that all minor comments will be implemented in the revised manuscript (including suggested references).  %Note that in what follows, all citations correspond to references of the original paper submission.

% EFFECT OF CORRECTED FOREARM ANNOTATIONS
\textbf{[R23] Are our annotations better? How the annotations affect the final performance?}
The only way to show that our annotations are more accurate than existing ones is by providing qualitative results and visual inspection (c.f.~Fig.7). In general, it is impossible to quantitatively compare the quality between different annotation sets, because there are no annotations that could be considered as ground truth.  Furthermore, in supervised learning better and more consistent annotations always lead to better models. An example of that is demonstrated in Fig.~A1, where the same algorithm, a cascade regression method, is learnt from noisy annotations (from crowd-sourcing) and from consistent annotations by our method. The quantitative experiments are performed on the widely used BBC Pose database because it has by far the most accurate annotations compared to the rest databases (FLIC, MPII etc.). As mentioned in the paper, we will provide all  corrected landmarks to the community upon acceptance.

% HUMAN EFFORT COMPARED TO HUMAN ANNOTATION
\textbf{[R23, R24] Human annotation effort.}
From our experience, the manual annotation of the body parts requires about 2 minutes per image. Thus, the manual annotation of FLIC database that includes about 20k images would require about 667 human working hours. On the contrary, our system requires the annotation of a few hundred images. The manual annotation of 700 such images took about 23 hours of human effort. Then these annotations were used by our pipeline to generate landmarks for all the FLIC images (and many other images). Additionally, according to an experiment we conducted, annotating images of faces and ears in terms of landmarks requires about 8.5 and 6.8 minutes respectively, while only 42 and 38 seconds are required to annotate them w.r.t.~curves. It becomes evident that the human effort required by our system is much smaller.

% building dense AAMs of patch-based AAMs with our pipeline cost 30 minutes for training with 600 images and fitting models cost the same time as normal AAM does. In our cases 5s for each fitting. Also, curve annotation requires much less effort, according to our experiment, putting landmarks on faces and ears requires 511s and 408s for each image including quality control, while only 42s and 38s need for annotating with curves.

% Manually annotating MPII (40K candidates), FLIC (20K+ candidates) require approximately 40K, 20K minutes correspondingly, while annotating using base line requires 700 minutes manually annotation + 40 minutes building time + 5s fitting time for each image. So 4K minutes and 2.3K minutes required for annotating entire MPII and FLIC dataset.

% FOCUS ON WRIST AND ELBOW
\textbf{[R23, R25] Body parts considered.} 
We  focus on the wrist and elbow because they are the most challenging body parts. R23 mentions that the hip and head are the most inconsistent body parts, but we disagree: Both LSP and MPII datasets report lower detection rates on wrist (MPII: 77\%, LSP: 79\%) and elbow (MPII: 84\%, LSP: 84\%), compared to hip (MPII: 83\%, LSP: 93\%) and head (MPII, LSP: 96\%)~[5]. In addition, the recently published system by Google (DeepPose) focuses on the elbow and wrist, which further demonstrates the challenges of these body parts and the interest they have attracted by the current state of the art. Also, Fig.6 demonstrates that elbow and wrist are the most noisily annotated points. Especially for the head, we believe it is very easy to accurately detect it, given the advances in head detection. In any case, our plan is to extend this work to include more parts (torso etc.), as well as a full body model, as suggested by R25.

% ARE OUR LANDMARKS CONSIDERED GROUNDTRUTH
%\textbf{[R23] Can our landmarks be considered as groundtruth?} As shown in our %qualitative results (figure 7), our landmarks are, in most cases, much more %accurate than the existing landmarks. Figure~\ref{fig:noise_fitting} shows some %indicative results of fitting a model trained on our landmarks and on noisy %landmarks (crow. The results are very indicative of the impact that more accurate 5annotations could have to the effectiveness of an SDM.

% INITIALISATION OF FOREARM FITTING
\textbf{[R23, R25] Initialisation of forearm fitting.}
It is initialised with our in-house implementation of a 
CNN pose estimation system, similar to~[41]. We remark that for a point-to-point error of 10, we get an improvement of about 10-15\% (in absolute terms, i.e. our method has 88\% at 10, while initilisation is at 73\%) for both wrist and elbow. In the final version of the paper, the initialisation CED curves will be included in Fig.10 (they are close to the yellow lines).
%Figure~\ref{fig:hand_benchmark} shows the results along with the initialisation curve, as requested.
%%
\begin{figure}[t!]
    \newcommand{\ofh}{0.24\columnwidth}
    \centering
    \includegraphics[height=\ofh]{face_all}
    \includegraphics[height=\ofh]{face2_all}
    \includegraphics[height=\ofh]{arm_all}
    \includegraphics[height=\ofh]{arm2_all}
    \caption*{Figure A1. Visualisation of fitting models (faces and forearms) trained with noisy annotation (blue dots) and trained with consistent annotation (green dots).}
    \label{fig:noise_fitting}
    \vspace{-14pt}
\end{figure

% LFPW EXPERIMENT
\textbf{[R23] Facial experiment on LFPW dataset.}
Unfortunately, LFPW database does not provide the actual images but only the original hyperlinks (from news-cites etc.) to them and currently most of them are inactive. We got the 224 from the authors of [24] who managed to download them 3 years ago. To keep it simple we have used state-of-the art AAMs such as [1,52] but on pixel intensities. State of the art includes the same methods but with image features such as HoGs [1,52] (in the revised paper, in Fig. 9, we will include AAMs on features).

% PERFORMANCE ANALYSIS OF PROPOSED METHOD'S COMPONENTS
\textbf{[R23, R24, R25] Performance analysis of each component is missing. Difficult to implement.} 
Unfortunately, due to limited space we could not show the effect of all parameters involved in the pipeline. We plan to make such an analysis on a journal extension of this work. Nevertheless, we are committed to make the whole pipeline publicly available. 
% , as done with all software developed in our group,

\textbf{[R25] Full body model.} Currently, our aim, starting with arms, is to create contour models for each body part. Then, we are thinking of creating an articulated model that connects each of them, but this is future work on the topic.

\textbf{[R25] Dimension of the $\mathbf{q}_i$ vector.} 
Every $\mathbf{q}_i$ is a sequence of 2D displacement vectors. It maps the image index $n$ ($n=1,..,N_t$ ) to a 2D vector 
$\mathbf{q}_i(n)$.  
Equivalently, $\mathbf{q}_i$ can be seen as a $2 N_t$ - dimensional vector, containing the displacements for all training images. 
%that correspond to the $i$-th basis element.

 %We plan to make such an analysis on a journal extension of this work. We refer the %Reviewer to the relevant bibliography of each component (SVS [39], NICP [3], Flow %[27], dAAM [2], patch AAM [52]).

%\begin{figure}[t!]
%    \centering
%    \includegraphics[width=0.48\columnwidth]{noise2}
%    \caption{Benchmark between AAM trained with consistant and noisy data}
%    \label{fig:noise}
%\end{figure}

% \textbf{[R23, R25] Effort of complete contour annotation of the forearm}. In the case of forearm, taking only two points into account actually requires slightly more effort. But for usecase that requires more points, our pipeline will be much faster. Moreover, contour annotation can work with semantic segmentation which can make the pipeline fully automatic (boundary of semantic segmentation gives contours needed).

% \textbf{[R23, R25] Manual effort spent on obtaining richer annotation worth ... SDM can be built with less effort does not seem to apply to the lower arm model ... }

%{\small
%\bibliographystyle{ieee}
%\bibliography{bib}
%}

\end{document}

### 2016-01-27 rebuttal/other_rebuttal.tex

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr_rebuttal}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

%%%%%%%%% PAPER ID  - PLEASE UPDATE
\def\cvprPaperID{89} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
\setlength{\parindent}{0pt}

\usepackage{caption}

\begin{document}

%%%%%%%%% TITLE - PLEASE UPDATE
\title{Dynamic Probabilistic Linear Discriminant Analysis}  % **** Enter the paper title here

\maketitle
\thispagestyle{empty}

%%%%%%%%% BODY TEXT - ENTER YOUR RESPONSE BELOW
We thank Reviewers (R) 10 and 22 for their valuable feedback. R10 ("A good paper") and R22 ("The proposed methods are clearly derived from sound mathematical principles") are positive regarding the paper. R13 strongly rejects the paper (being very confident), even though it is evident that he/she is not knowledgeable at all regarding the field (probabilistic machine learning). We kindly request the assistance of the Area Chair on that.    %Note that in what follows, all citations correspond to references of the original paper submission.

\textbf{[R10] Regarding Typos. }. Since, submission we have thoroughly edited the text to spot and correct all typos. An extended supplementary material will also accompany the paper (may it be accepted).

\textbf{[R10] The authors should provide some experiments regarding the training time of the model}. Since, the methods scale linearly with regards to number of videos, frames, people etc. it is quite fast to train and test them. For example for the largest training set it takes few minutes to train (not more than 10 mins using un-optimised Matlab code). Testing, if carefully implemented, can be performed real-time. Detailed complexity evaluations will be included for each individual experimental setup.

\textbf{[R10] The authors should consider comparisons of the proposed model with some other previous works like [20], [21].}. [20] is a CCA model that cannot be used for classification but for identifying shared structures. [21] proposes a unified view on probabilistic techniques. The only technique form [21] that is relevant is an alternative static PLDA which will include in the revised manuscript (performance very close to the tested static PLDA).

% EFFECT OF CORRECTED FOREARM ANNOTATIONS
\textbf{[R13] How correlation are found if $\mathbf{A}=0.98\mathbf{I}$? If $A_{ij}$ very close to 1, how to differentiate $W_{ij}^t$ and $W_{ij}^{t+1}$? In Eq.(19), how to define the value of $A_{ij}$?}
First we would like to explain that $\mathbf{A}=0.98\mathbf{I}$ is used ONLY to PRODUCE artificial data in section 4.1 (a legitimate choice though, since it corresponds to a Slow feature dynamical system \cite{turner2007maximum}). In all experiments $\mathbf{A}_{ij}$ is found from the data through the application of the EM algorithm (standard practise, see 444-454). We refer the reviewer to standard textbooks \cite{bishop2006pattern} and tutorial papers \cite{roweis1999unifying} for further details on Linear Dynamical Systems (LDS).

% HUMAN EFFORT COMPARED TO HUMAN ANNOTATION
\textbf{[R13] In Eq.(1), what the meaning of $\mathbf{y}_i$? There is no explanation in Section 2.1.. In Eq.(7), what is $Dy$ means? There is also no explanation in this section. etc.}
There is a clear explanation in 181-184 of the original paper. Unfortunately, we do not have space to describe in full detail standard models such as PPCA. Again we refer R13 to  \cite{bishop2006pattern,roweis1999unifying}. Equation (7) is a condensed form of equation (6) (it is easy to make the association between variables).

% building dense AAMs of patch-based AAMs with our pipeline cost 30 minutes for training with 600 images and fitting models cost the same time as normal AAM does. In our cases 5s for each fitting. Also, curve annotation requires much less effort, according to our experiment, putting landmarks on faces and ears requires 511s and 408s for each image including quality control, while only 42s and 38s need for annotating with curves.

% Manually annotating MPII (40K candidates), FLIC (20K+ candidates) require approximately 40K, 20K minutes correspondingly, while annotating using base line requires 700 minutes manually annotation + 40 minutes building time + 5s fitting time for each image. So 4K minutes and 2.3K minutes required for annotating entire MPII and FLIC dataset.

% FOCUS ON WRIST AND ELBOW
\textbf{[R13] Some state-of-the-art methods for YouTube face database, FERA database should also be compared in this section.} 
Comparisons exist in the original paper (please see 691-701 and 782-786). We will summarise them in a Table.

% ARE OUR LANDMARKS CONSIDERED GROUNDTRUTH
%\textbf{[R23] Can our landmarks be considered as groundtruth?} As shown in our %qualitative results (figure 7), our landmarks are, in most cases, much more %accurate than the existing landmarks. Figure~\ref{fig:noise_fitting} shows some %indicative results of fitting a model trained on our landmarks and on noisy %landmarks (crow. The results are very indicative of the impact that more accurate 5annotations could have to the effectiveness of an SDM.

% INITIALISATION OF FOREARM FITTING
\textbf{[R22]  The authors do not compare their techniques to any of the competing techniques that could be trained on the same data sets such as LBP, Convolutional Neural Networks or Support Vector Machines using Gabor filters or Gaussian Derivatives.}
The aim of the experiments were to show that our general machine learning methodology performs close to state-of-the-art even though combined with very simple features (IGO features).  Nevertheless, we compare with state-of-the-art techniques in 691-701 that use elaborate features.
Since, our method is a generic machine learning tool it can be used on top of arbitrary features (e.g. LBP and Gabor filters). For example we have now performed experiments using HoG and LBP features and the results have improved a lot.

% LFPW EXPERIMENT
\textbf{[R23] Deep learning versus probabilistic methods.}
We agree with R23 that when a lot of data are available deep learning techniques can provide a good alternative. This is true for face recognition. But for facial expression recognition (FERA experiments) to the best of our knowledge there is no competing deep network proposed yet. We would like to note that DCNNs provide mainly good features which can be combined with any classification technique (including ours).

% PERFORMANCE ANALYSIS OF PROPOSED METHOD'S COMPONENTS
\textbf{[R22] There are very large data sets , such as the CASIA Webface data set (10,575 subjects and 494,414 images), that are available for deep learning techniques.} 
We agree that CASIA database is a valuable source of data (we will cite CASIA in the revised version). Nevertheless, it does not provide the amount of data necessary to match the performance of FaceNet by Google (200 Million) or DeepFace by FaceBook.

% , as done with all software developed in our group,

\textbf{[R10-R25] Description of state-of-the-art (R25) and a table (R10).} We will include a table in the revised paper summarising all the state-of-the-art techniques, as well as further description in the text.

Finally, it goes without saying that all minor comments will be addressed. Also, we have already made steps towards making the code publicly available.

 %We plan to make such an analysis on a journal extension of this work. We refer the %Reviewer to the relevant bibliography of each component (SVS [39], NICP [3], Flow %[27], dAAM [2], patch AAM [52]).

%\begin{figure}[t!]
%    \centering
%    \includegraphics[width=0.48\columnwidth]{noise2}
%    \caption{Benchmark between AAM trained with consistant and noisy data}
%    \label{fig:noise}
%\end{figure}

% \textbf{[R23, R25] Effort of complete contour annotation of the forearm}. In the case of forearm, taking only two points into account actually requires slightly more effort. But for usecase that requires more points, our pipeline will be much faster. Moreover, contour annotation can work with semantic segmentation which can make the pipeline fully automatic (boundary of semantic segmentation gives contours needed).

% \textbf{[R23, R25] Manual effort spent on obtaining richer annotation worth ... SDM can be built with less effort does not seem to apply to the lower arm model ... }

{\small
\bibliographystyle{ieee}
\bibliography{egbib}
}

\end{document}

### 2016-04-06 contents/introduction.tex
Updates from ShareLaTeX
Rebuttal Final

In this paper, we propose a solution for annotating an object with regards to its deformations that requires considerably less effort compared to manual annotation and, at the same time, can be used to define statistical deformations for objects without a consistent set of landmarks. We employ the proposed method in order to construct SDMs based on the outline of human body parts (i.e., arms and legs). To the best of our knowledge, this is the first time that an SDM is proposed to be trained on the outline of human body parts and not on their skeleton joints. The proposed SDM can also be used to provide accurate and consistent annotations for several of the body joints (such as wrist, elbow etc.).

To this end, we argue and empirically demonstrate that it is better to annotate an object with regards to a set of continuous lines that describe its shape. An example is provided in Figure \ref{fig:intro}, which compares the standard landmark annotations that are employed in the current literature with the proposed curve annotations for arms, ears and faces. It becomes evident that the curve annotations avoid the inherent ambiguity of placing sparse landmarks and offer a richer description of the object's shape.
%
%where the shape of the human arm is annotated with continuous lines instead of a number of discrete landmarks. Similarly, the shape annotations for an ear and a face are shown in the same figure. Annotating images by following this procedure takes 42, 38 and 43 secs per image for arm, faces, ears correspondingly.
Furthermore, these curves can be automatically generated by recently proposed methods that perform discriminative segmentation of objects \cite{luo2013pedestrian,liu2015matching}.

Furthermore, we capitalise on recent advances on multiframe optical flow estimation \cite{Garg:2013hu,tomasi2012dense,snape15faceflow} and show that the relevant methodologies have matured enough to densely annotate the proposed shapes using either simplistic or even more sophisticated and robust shape representation methods \cite{Nguyen2013}.
In particular, in order to build dense correspondences between different shape instances of the same object class, we jointly estimate the optical flow among all the instances by imposing low-rank constrains, an approach that we call \emph{Shape Flow}. Multiframe optical flow has originally been applied on video sequences, relying on the assumptions of colour consistency and motion smoothness \cite{Garg:2013hu}. However, these assumptions do not hold in our case, where we have a collection of shapes. Therefore, we introduce appropriate modifications based on the consistency of image-based shape representation, as well as low rank priors.
%
%On the other hand,  while shapes having no motion smoothness completely as all training data are independent. Low rank constrain plays important role to simulate motion smoothness assumption.

@@

%To the best of our knowledge this is the first methodology which can %create a dense statistical model for AAM which can operate in %"in-the-wild".

%[..]. There are methods such as [..] and [..] but ...

In summary, the contributions of this paper are the following:
\begin{itemize}

@@

  \item We report state-of-the-art quantitative and qualitative results on the task of human body parts localisation by employing a patch-based SDM trained on the outline landmarks that are sampled by the dense correspondences. Our proposed model outperforms all current state-of-the-art techniques that are trained on skeleton joints

>>>>

In this paper, we propose a solution for annotating an object with regards to its deformations that requires considerably less effort compared to manual annotation and, at the same time, can be used to define statistical deformations for objects without a consistent set of landmarks. We employ the proposed method in order to construct SDMs based on the outline of human body parts (i.e., arms and legs). While majority of proposed SDMs for human poses are trained on skeleton joints, training SDMs based on outline of human body parts have shown remarkable advantages~\cite{zuffi2012pictorial} (?). The proposed SDM can also be used to provide accurate and consistent annotations for several of the body joints (such as wrist, elbow etc.).
To this end, we argue and empirically demonstrate that it is better to annotate an object with regards to a set of continuous lines that describe its shape. An example is provided in Figure \ref{fig:intro}, which compares the standard landmark annotations that are employed in the current literature with the proposed curve annotations for arms, ears and faces. It becomes evident that the curve annotations avoid the inherent ambiguity of placing sparse landmarks and offer a richer description of the object's shape.

Furthermore, these curves can be automatically generated by recently proposed methods that perform discriminative segmentation of objects \cite{luo2013pedestrian,liu2015matching}.

Furthermore, we capitalise on recent advances on multiframe optical flow estimation \cite{Garg:2013hu,tomasi2012dense,snape15faceflow} and show that the relevant methodologies have matured enough to densely annotate the proposed shapes using either simplistic or even more sophisticated and robust shape representation methods \cite{Nguyen2013}.
In particular, in order to build dense correspondences between different shape instances of the same object class, we jointly estimate the optical flow among all the instances by imposing low-rank constrains, an approach that we call \emph{Shape Flow}. Multiframe optical flow has originally been applied on video sequences, relying on the assumptions of colour consistency and motion smoothness \cite{Garg:2013hu}. However, these assumptions do not hold in our case, where we have a collection of shapes. Therefore, we introduce appropriate modifications based on the consistency of image-based shape representation, as well as low rank priors.

@@

In summary, the contributions of this paper are:
\begin{itemize}

@@

  \item We report state-of-the-art quantitative and qualitative results on the task of human body parts localisation by employing a patch-based SDM trained on the outline landmarks that are sampled by the dense correspondences. Our proposed model outperforms all current state-of-the-art techniques that are trained on skeleton joints.
